Language learning in context via recursive pragmatic reasoning

### 2013-05-31 praglearn.tex
missing space
tewaks

\begin{table}[t]
\label{tab:results}
\begin{center}
\begin{tabular}{lccccc}
\toprule
Phenomenon & Ref. & WL & PI & PI+U & PI+WL \\
\midrule
Interpreting specificity implicature & \cite{grice1975} &  & x &x & x\\
Interpreting Horn implicature & \cite{horn1984} & &  & x& x\\
Cross-situational learning & \cite{smith2008} & x& & & x \\
Disambiguating new words using old words & \cite{markman1988} &x  & & & x \\
Disambiguation without mapping & \cite{horst2008} &  & & & x \\
Emergence of novel \& efficient communicative systems & \cite{galantucci2005} & & & & x \\
Lexicalization of Horn implicature & \cite{horn1984} & & & & x \\
Non-lexicalization of specificity implicature & \cite{levinson2000} & & & & x \\
\hline
\end{tabular}
\end{center}
\caption{Empirical results and references. WL refers to the word
  learning model of \cite{frank2009}; PI refers to the recursive
  pragmatic inference model of \cite{frank2012}; PI+U refers to the
  pragmatic inference model of \cite{bergen2012} which includes
  lexical uncertainty, marginalizes it out, and then recurses. Our
  current model is referred to here as PI+WL, and combines pragmatic
  inference with word learning.}
\end{table}

@@

Humans, however, unlike traditional game theoretic models, do make the
inference that if two utterances seem equivalent except that one is
more costly to utter, then the costly utterance refers to an unusual
event. We call this pattern \textit{Horn implicature}, after
\cite{horn1984}. For instance, ``Lee got the car to stop'' implies
that Lee used an unusual method (e.g.~not the brakes) because, had he
used the brakes, the speaker would have chosen the simpler and shorter
(less costly) expression, ``Lee stopped the car''
\cite{horn1984}. Surprisingly, Bergen et al \cite{bergen2012} show
that the key to achieving this favorable result is ignorance. If a
listener assigns equal probability to her partner using the good
system or the bad system, then their best bet is to estimate
$P_S(\word | \obj)$ as the average of $P_S(\word | \obj, \text{good
  system})$ and $P_S(\word | \obj, \text{bad system})$. These might
seem to cancel out, but in fact they do not. In the good system, the
speaker has higher expected utility than the bad system; therefore, a
soft-max agent in the bad system has noiser behavior than in the good
system, and their behavior in the good system dominates the average,
breaking the symmetry. Similar reasoning applies to an uncertain
speaker.

@@

@@

\section{The role of pragmatics in language acquisition}

@@

\section{Mutual learning situations}

@@

Language learners and language users must consider word meanings both
within and across contexts. A critical part of this process is
reasoning pragmatically about agents' goals in individual
situations. In the current work we assume that agents communicating
with one another assume that there is a shared conventional lexicon
which they both rely on. They then
reason recursively about how this lexicon should be used to convey
particular meanings in context. This pair of assumptions allows us to
create a model that unifies two previously separate strands
of modeling work on language usage and acquisition, and account for a
variety of new phenomena. In particular, we consider new
explanations of disambiguation in early word learning and the acquisition of
scalar adjectives, and demonstrate that our model is capable of developing
novel and efficient communicative systems through iterated learning
within the context of a single simulated conversation.

>>>>

\begin{table}[t]
\label{tab:results}
\begin{center}
\begin{tabular}{lccccc}
\toprule
Phenomenon & Ref. & WL & PI & PI+U & PI+WL \\
\midrule
Interpreting specificity implicature & \cite{grice1975} &  & x &x & x\\
Interpreting Horn implicature & \cite{horn1984} & &  & x& x\\
Disambiguating new words using old words & \cite{markman1988} &x  & & & x \\
Disambiguation without mapping & \cite{horst2008} &  & & & x \\
Emergence of novel \& efficient communicative systems & \cite{galantucci2005} & & & & x \\
Lexicalization of Horn implicature & \cite{horn1984} & & & & x \\
Non-lexicalization of specificity implicature & \cite{levinson2000} & & & & x \\
\hline
\end{tabular}
\end{center}
\caption{Empirical results and references. WL refers to the word
  learning model of \cite{frank2009}; PI refers to the recursive
  pragmatic inference model of \cite{frank2012}; PI+U refers to the
  pragmatic inference model of \cite{bergen2012} which includes
  lexical uncertainty, marginalizes it out, and then recurses. Our
  current model is referred to here as PI+WL, and combines pragmatic
  inference with word learning.}
\end{table}

@@

Humans, however, unlike traditional game theoretic models, do make the
inference that if two utterances seem equivalent except that one is
more costly to utter, then the costly utterance refers to an unusual
event. We call this pattern \textit{Horn implicature}, after
\cite{horn1984}. For instance, ``Lee got the car to stop'' implies
that Lee used an unusual method (e.g.~not the brakes) because, had he
used the brakes, the speaker would have chosen the simpler and shorter
(less costly) expression, ``Lee stopped the car''
\cite{horn1984}. Surprisingly, Bergen et al. \cite{bergen2012} show
that the key to achieving this favorable result is ignorance. If a
listener assigns equal probability to her partner using the good
system or the bad system, then their best bet is to estimate
$P_S(\word | \obj)$ as the average of $P_S(\word | \obj, \text{good
  system})$ and $P_S(\word | \obj, \text{bad system})$. These might
seem to cancel out, but in fact they do not. In the good system, the
speaker has higher expected utility than the bad system; therefore, a
soft-max agent in the bad system has noiser behavior than in the good
system, and their behavior in the good system dominates the average,
breaking the symmetry. Similar reasoning applies to an uncertain
speaker.

@@

%% NOTE FOR REVISION: THIS BIT HERE ABOUT BERGEN MAKES NOAH SAD AND SHOULD BE RECONSIDERED.

@@

\section{Pragmatics in learning from a knowledgable speaker}

@@

\section{Pragmatic reasoning in the absence of conventional meanings}

@@

Language learners and language users must consider word meanings both
within and across contexts. A critical part of this process is
reasoning pragmatically about agents' goals in individual
situations. In the current work we treat agents communicating
with one another as assuming that there is a shared conventional lexicon
which they both rely on. They then
reason recursively about how this lexicon should be used to convey
particular meanings in context. This pair of assumptions allows us to
create a model that unifies two previously separate strands
of modeling work on language usage and acquisition and account for a
variety of new phenomena. In particular, we consider new
explanations of disambiguation in early word learning and the acquisition of
scalar adjectives, and demonstrate that our model is capable of developing
novel and efficient communicative systems through iterated learning
within the context of a single simulated conversation.
